Multi-level generalizations of collective spin dynamics and squeezing

### 2019-09-25 sun_squeezing/sun_notes.tex
simplified derivation for changing operator bases in collective interactions

There are two types of terms in the Hamiltonian in
\eqref{eq:H_int_spin}: two-body interactions and self-interactions,
respectively of the form
\begin{align}
  \sum_j X_j^\dag \otimes X_j
  = \sum_j \oket{X_j^\dag}\otimes\oket{X_j},
  &&
  \sum_j X_j^\dag X_j = \sum_j \oket{X_j^\dag X_j}.
\end{align}
Choosing a different basis $\set{Y_j}$ for $\B_n$ with norm $\N_Y$, by
resolution of the identity we find that
\begin{align}
  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
  = \f1{\N_Y^2} \sum_{j,k,\ell}
  \oket{Y_k^\dag} \obk{Y_k^\dag|X_j^\dag}
  \otimes \oket{Y_\ell} \obk{Y_\ell|X_j},
\end{align}
where
\begin{align}
  \obk{Y_k^\dag|X_j^\dag}
  = \tr\p{Y_k X_j^\dag}
  = \tr\p{X_j^\dag Y_k}
  = \obk{X_j|Y_k},
\end{align}
which allows us to simplify
\begin{align}
  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
  = \f1{\N_Y^2} \sum_{k,\ell}
  \oket{Y_k^\dag} \otimes \oket{Y_\ell}
  \sum_j \obk{Y_\ell|X_j} \obk{X_j|Y_k}
  = \f{\N_X}{\N_Y} \sum_k
  \oket{Y_k^\dag} \otimes \oket{Y_k}.
\end{align}
For the single-body self-interactions, we can similarly expand
\begin{align}
  X_j = \f1{\N_Y} \sum_k Y_k \obk{Y_k|X_j},
  &&
  X_j^\dag = \f1{\N_Y} \sum_k Y_k^\dag \obk{Y_k^\dag|X_j^\dag}
  = \f1{\N_Y} \sum_k Y_k^\dag \obk{X_j|Y_k},
\end{align}
which implies
\begin{align}
  \sum_j X_j^\dag X_j
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k^\dag Y_\ell
  \obk{X_j|Y_k} \obk{Y_\ell|X_j}
  = \f{\N_X}{\N_Y} \sum_k Y_k^\dag Y_k.
\end{align}
The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
takes an identical form in any orthonormal basis for the space $\B_n$
of operators on the Hilbert space of an $n$-level system, up to
appropriate re-scaling factors $\N_X/\N_Y$.

>>>>

The collective interaction Hamiltonian in \eqref{eq:H_int_spin} takes
the form
\begin{align}
  H = \sum_{p,q,j} {X_j^{(p)}}^\dag X_j^{(q)},
\end{align}
which essentially consists of two-body interactions ($p\ne q$) and
self-interactions ($p=q$), respectively of the form
\begin{align}
  \sum_j X_j \otimes X_j^\dag,
  &&
  \sum_j X_j X_j^\dag.
\end{align}
Choosing a different basis $\set{Y_j}$ for $\B_n$ with corresponding
norm $\N_Y$, we can expand
\begin{align}
  X_j = \f1{\N_Y} \sum_k Y_k \obk{Y_k|X_j},
  &&
  X_j^\dag = \f1{\N_Y} \sum_k Y_k^\dag \obk{Y_k^\dag|X_j^\dag}
  = \f1{\N_Y} \sum_k Y_k^\dag \obk{X_j|Y_k},
\end{align}
where we used the fact that
\begin{align}
  \obk{Y_k^\dag|X_j^\dag}
  = \tr\p{Y_k X_j^\dag}
  = \tr\p{X_j^\dag Y_k}
  = \obk{X_j|Y_k}.
\end{align}
By resolution of the identity, we can therefore find
\begin{align}
  \sum_j X_j \otimes X_j^\dag
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k \obk{Y_k|X_j}
  \otimes Y_\ell^\dag \obk{X_j|Y_\ell}
  = \f{\N_X}{\N_Y^2} \sum_{k,\ell} Y_k \otimes Y_\ell^\dag
  \obk{Y_k|\I|Y_\ell}
  = \f{\N_X}{\N_Y} \sum_k Y_k \otimes Y_k^\dag,
\end{align}
and similarly
\begin{align}
  \sum_j X_j X_j^\dag
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k Y_\ell^\dag
  \obk{Y_k|X_j} \obk{X_j|Y_\ell}
  = \f{\N_X}{\N_Y} \sum_k Y_k Y_k^\dag.
\end{align}
The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
takes an identical form in any orthonormal basis for the space $\B_n$
of operators on the Hilbert space of an $n$-level system, up to
appropriate re-scaling factors $\N_X/\N_Y$.
